---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# annotater

The goal of annotater is to annotate package load calls in strings and R scripts, so we can have an idea of the overall purpose of the libraries we're loading.

This project came about after teaching workshops or helping peers and realizing that many issues relate to package installation failures and dependency issues for packages that are not even used in a problematic script. Scripts get passed around, code is copied and pasted, and we might not know what certain packages are for.


## Installation

You can install the development version of annotater from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("luisDVA/annotater")
```
Alternatively:

``` r
# install.packages("remotes")
remotes::install_github("luisDVA/annotater")
```

## Example

This is a basic example with a simple character string.

```{r example}
library(annotater)
test_string <-c("library(boot)\nrequire(Matrix)")
writeLines(annotate_pckg_calls(test_string))
```

 Entire .R files can also be parsed and annotated with the _annotate_script_ function.
 
 
_annotater_ now includes an RStudio addin to make  annotations on the active R file in the source pane.

Try it out!
